---
title: "ForageFishReport"
author: "Jaydi Swanson"
date: "12/15/2021"
output: html_document
---
##INTRODUCTION##

As global populations continue to increase, so does reliance on increasingly stressed and limited marine resources. Fisheries are a major source of food and economic growth for many countries around the world, so understanding how to manage them sustainably is of critical importance. Historically, fisheries management has focused on single stocks and maximizing sustainable catch based on parametric models for individual species. This, however, ignores the many complex biological and environmental ecosystem interactions that inform population dynamics and underlie variation. There is a growing push towards modelling and management strategies that account for ecosystem-based interactions rather than analyzing single stocks in isolation (Howell et al, 2021). Ecosystem models can often be complex, however, and focus on long term trends which may be unhelpful for stock management in the short term. Thus, we try to simplify the problem by focusing on one interaction, that between forage fish and their piscivorous predators.

Forage fish support a majority of the ocean's fisheries, either as the target species or the main food source for other commercially-relevant fish (Pikitch et al, 2012). Understanding the relationship between forage fish and their predators is imperative for fisheries management as overfishing either the prey or the predator species can have a major impact on the stability of the fishery (Howell et al, 2021). Current fisheries management solutions use catch-based data and single species stock assessments to quantify the populations of the target fish species. These models do well by giving important information about the current state of the stock; however, they often overestimate the extent to which a stock is being overfished or collapsed (Branch et al, 2010). An issue with using catch-based and single-species assessments for forage fish is that the forage fish are more interdependent and have more pressures than other commercial fish species. Forage fish make up a large proportion of the global catch, with five of the top ten fished species being forage fish, which keeps them in a constant state of anthropogenic pressure (Pikitch et al, 2012). Along with heavy fishing pressure, forage fish make up a large portion of the diets of many commercial fish as well as marine mammals and seabirds, creating the other source of pressure and highlighting their importance in the ecosystem.  

The goal of this project was to determine if a simple two-species model could be used to achieve more desirable outcomes for predator and forage fish yields than similar single-species models. The models used were single species and two species Ricker models, and an Empirical Dynamic Modelling (EDM) approach using convergent cross-mapping was employed to determine if machine-learning methods could similarly validate the presence of a trade-off relationship between forage fish and their predators. If the two species model produces higher yields than the single species model this will be evidence that  ecosystem based management is more effective than the single species stock assessment based management. The basis for this hypothesis is that the two species model is closer to ecosystem based management as it takes into account the interactions between two species rather than looking at the dynamics of a single species. Since the goal of all fisheries is to optimize yield, stockholders and management want to find the models that will allow them to harvest at the optimal level, which we seek to prove can coincide with fishing at rates that are most sustainable to the ecosystem as a whole. 


```{r message = FALSE}
#Load ggplot2
library(ggplot2)
library(rEDM)
```


```{r include = FALSE}
#Loading all files
fname_EDM_results = "./Data/EDM_results.Rdata"

source("./Functions/PopulationEquations.R")
source("./Functions/simulateData.R")
source("./Functions/TradeOffAnalysis.R")
source("./Functions/Plots.R")
source("./Functions/MSY.R")
source("./Functions/CCM.R")
source("./Data/PlotsToLoad.R")


if(!file.exists(fname_EDM_results)){
source("./Data/PlotsAndData.R")
  save(cVSccm_no0,cVSccm,file=fname_EDM_results)
}else{
  load(fname_EDM_results)
}

```


##TRADE-OFF ANALYSIS METHODS##

###Perretti Model

To accomplish the first goal of our project, showing how ecosystem based management is more sustainable than the currently used single species models, we sought to prove that there is a trade off relationship between forage fish species and their predators. Validation for this trade-off supports the idea that species should not be managed without consideration for the species they impact or are impacted by. To do this, we first needed to create a two-species model that acts as a proxy for ecosystem based management. This model would demonstrate interactions between the forage fish species and predator species and would show the impacts of forage fish dynamics on predator yields and biomass. 

In beginning to refine our model, we created simulations of trade off analysis using the two-species model presented in Perretti et al. (cite). This model is relatively simple, but it does introduce a coupling strength factor between the two species. Through a series of functions, we simulated multiple time series with varying fishing mortality on both species. Predator and prey biomass and yields  were averaged for each time series and plotted against the corresponding fishing effort to show contours of biomass. Our initial parameters were taken from the supplemental information from Perretti et al., though we adjusted the carrying capacities based off of the common rule of 10% energy transfer and biomass size up a trophic level, which means the striped bass carrying capacity should be about 10% of the menhaden's.

#The functions can be found in Functions/PopulationEquations.R, Functions/simulateData.R, and Functions/TradeOffAnalysis.R. The functions were plotted through a function in Functions/PlottingThings.Rmd. Parameters for this data set can be found in Data/dataf.R
```{r}
#plots of Perretti model data
plotTradeOff(Perretti_sim)
```
In the first graph, "Trade Off Contour Analysis-Predator Biomass", total predator biomass is plotted against both predator and forage fish fishing mortality. According to this model, the contours do not change with forage fish fishing mortality until unrealistically high fishing mortality is reached, which suggests that there is no relationship between forage fish fishing dynamics and predator biomass at real world fishing efforts. 

The following two graphs demonstrate predator yield and forage fish yield with changing fishing dynamics, and they show similar results as the yields only change with the specified species fishing mortality at realistic fishing levels. 

A limitation with the Perretti et al. two-species model is the chance for populations to become negative. The Perretti model determines population size before coupling and fishing effects with the following equation, where x0 is the initial population size, r is intrinsic growth rate, and K is carrying capacity: $ x^2 $

#Full models can be found in ./Functions/PopulationEquations.R
$$
x1 = x0*r*(1-(x0/K)) 
$$
If K is less than x0, the population becomes negative. Because this is not possible in the real world, we chose to abandon the Perretti model and search for a more realistic one. The trade-off relationship is very weakly demonstrated in this approach, so with a better two-species model, we can demonstrate a stronger trade-off relationship.

###Ricker Model

Our search for a more accurate model led us to the Ricker single-species model, which we adapted for a two-species approach by adding the same coupling factor presented in Perretti et al..

```{r}
#Models can be found in ./Functions/PopulationEquations.R

forageFishRicker_noise <- function(x0, y0, r, K, c, f){
  
  #population without fishing
  x1 = ((x0*exp(r*(1-(x0/K)))) - (c*x0*y0))*rlnorm(1, meanlog=0, sdlog=0.005)
  #population after fishing
  x1f = x1 - (f*x1)
  
  return(x1f) #returns population next year after fishing
}


predatorRicker_noise <- function(y0, x0, r, K, c, f){
  
  #population without fishing
  y1 = ((y0*exp(r*(1-(y0/K)))) + ((c*x0*y0)))*rlnorm(1, meanlog=0, sdlog=0.005)
  #population after fishing
  y1f = y1 - (f*y1)
  
  return(y1f) #returns population next year after fishing
}
```

To test this model and see if it worked correctly for our chosen species, Atlantic menhaden and striped bass, we searched for starting parameters in the literature. Though we were able to determine information on the relationship between growth rates and a relative coupling factor based on the proportion of striped bass diet made up of menhaden, these parameters did not produce realistic data with our model. 

Final parameters for the Ricker two-species model were chosen arbitrarily to optimize the output of our model. The 10% energy transfer rule was maintained by keeping the striped bass carrying capacity and initial population size at 10% of the Atlantic menhaden population's parameters. The striped bass growth rate was also kept at less than half of the menhaden growth rate to simulate realistic growth. All coupling values produced a trade-off relationship, leading us to set c to 0.2 in an attempt to remain close to the value presented in Perretti et al., though we will compare higher and lower coupling strengths. 

The following parameters were used for the Ricker two species approach:
```{r include = FALSE}
x0 <- 0.5 #Initial forage fish population size
  
y0 <- 0.05 #Initial predator population size
  
rx <- 2.0 #Forage fish intrinsic growth rate
  
ry <- 0.78 #Predator intrinsic growth rate
  
Kx <- 1.0 #Forage fish carrying capacity
  
Ky <- 0.1 #Predator carrying capacity
  
cx <- 0.2 #Forage fish coupling strength
  
cy <- 0.2 #Predator coupling strength

numTimeSteps <- 50 #100 time steps in each simulation 

numMeanYears <- 50 #averaging the last 50 years of each
```

After determining our parameters, we produced the same graphs as shown above in the Perretti methods, this time using our generateTradeOff_Ricker function (./Functions/TradeOffAnalysis.R). This function takes an input of the parameters listed above along with vectors of fishing mortality to loop over and a number of times for the simulation to be ran. Time series at each fishing mortality pair are created inside the loop using a function called simulateRicker_noise (./Functions/simulateData.R), which outputs a data frame containing a single time series of predator and forage fish population sizes over time at constant fishing mortality. Then, biomass is averaged across the time series for both species and inputted into a new data frame in generateTradeOff_Ricker. Yields are also calculated for each series. The final output is a data frame with fishing mortality pairs, predator biomass and yield, and forage fish biomass and yield for a number of simulations. 

For the following analysis, we used a single data set from the above parameters with fishing mortality from 0.0 to 1.0 and ### simulations. 

```{r}
#Variable toRicker_lowC_data is a data frame with data for a Ricker two-species model simulation with the above parameters
head(toRicker_lowC_data) 
```

Once the data is simulated, we first plot the population dynamics to ensure the model is producing reasonable data. 

```{r}
#Dynamics with no fishing mortality (fx = 0, fy = 0)
tsRicker_noFishing_plots #./Functions/PlottingThings.Rmd
```

```{r}
#Dynamics with fishing mortality (fx = 0.5, fy = 0.3)
tsRicker_Fishing_plots #./Functions/PlottingThings.Rmd
```
With and without fishing, both populations fluctuate around an equilibrium, though the equilibrium are lower with fishing.  In the trade off graphs without fishing, we see that predator population size decreases some as forage fish population size increases. We assume that the increase in forage fish follows a decrease in predator biomass. When fishing is applied, striped bass biomass is  higher when forage fish biomass is highest.

After verifying that our populations are fluctuating realistically with the chosen parameters, we plotted the trade-off relationships of biomass and yield. To compare our two-species (ecosystem based) model to a single species model, we simulated one time series of predator and forage fish biomass with a single species Ricker model using similar parameters, but no fishing mortality or coupling strength are added to the model. A single-species model does not account for dependency on other species or the ecosystem, so it often overestimates a maximum sustainable yield. When compared to two-species modeled data, single-species MSYs fall in regions that actually obtain lower yields and will cause decreases in biomass of both species. To demonstrate this, we overlaid our calculated single-species fishing mortality at MSY for both species on our two-species trade off graphs. This way, a visual representation of the discrepancies between the models is created. 


##RESULTS##
```{r}
#Graphs with coupling strength at 0.2

to_lowC_PBiomass
to_lowC_PYield 
to_lowC_FFYield
```
Figures: Solid red lines represent MSYs suggested by the single-species Ricker model. Dashed lines represent the maximum yield given by our two-species Ricker model. 

At a coupling strength of 0.2, predator biomass is highest at lower forager fish fishing mortality. The negative slant of the contours shows that predator biomass is impacted by forage fish dynamics. Similar trends are seen in the contours of predator and forage fish yields. Predator yield is highest when forage fish fishing mortality is low. This visualization validates the presence of a trade-off between Atlantic menhaden and striped bass. 

Because our analysis showed this trade-off relationship, we assumed the single-species models would not accurately determine sustainable or maximum yields. The intersection of the solid red lines falls outside of this maximum yield range. Though the dashed lines, the maximum yields from our two-species model, also fall outside of this range, they are much lower than the single-species suggestions. Our yields are also not representative of MSY because they are not accounting for a target biomass. However, this graph demonstrates that single-species management will produce lower yields by overestimating the MSY of both species. 

```{r}
#Graphs with coupling strength at 0.5

to_highC_PBiomass
to_highC_PYield
to_highC_FFYield
```
When coupling strength is increased to 0.5, we see a stronger trade-off relationship with even higher predator yields and biomass at lower forage fish fishing mortality.

#Results discussion for this portion of project

Through these methods, the existence of a trade-off relationship between forage fish species and their predators is verified. This information supports the growing motion for ecosystem based management as the effects of one species on the other are undeniable, especially in a predator-prey relationship. Still, there are questions that arise about our ability to quantify and detect these trade-off relationships to allow for ecosystem based fishing mortality suggestions. 

Specifically, machine learning can be used to detect causal relationships between variables. Empirical dynamical modeling is a nonparametric approach that models changes in nonlinear dynamic systems, such as populations over time. We wanted to test whether or not this method can detect trade-off relationships between species. 

##EMPIRICAL DYNAMIC MODELING METHODS##

```{r}
#Non-linear population dynamics plots

tsRicker_nl_plots
```
The dynamics created by the non-linear models are much more chaotic. As before, they fluctuate around the equilibrium point, though the fluctuations are wider. 

```{r}
ggplot(data = cVSccm, aes(x = c, y = predictability)) +
  geom_point(size = 2) +
  geom_smooth(data =cVSccm_no0, aes(x = c, y = predictability), method="lm", formula = y~log(x), fill="blue", fullrange=TRUE) +
  theme_bw() + 
  theme(plot.title = element_text(face="bold")) +
  labs(x = "Coupling Factor", y = "Predictability", 
       title = "Linear Model Coupling Factor vs \n Non-Parametric Convergent Cross Mapping Predictability")
  
```

