---
title: "ForageFishReport"
author: "Jaydi Swanson"
date: "12/15/2021"
output: html_document
---

Here, add intro stuff!


```{r}
#Loading all files

source("./Functions/PopulationEquations.R")
source("./Functions/simulateData.R")
source("./Functions/TradeOffAnalysis.R")
source("./Functions/PlottingThings.Rmd")
```

##Methods
###Perretti

First, we needed to create a two-species model, which acts as a proxy for ecosystem based management. This model would demonstrate interactions between the forage fish species and predator species. 

In beginning to refine our model, we created simulations of trade off analysis using the two-species model presented in Perretti et al. (cite). This model is relatively simple, but it does introduce a coupling strength factor between the two species. Through a series of functions, we simulated time series with varying fishing mortalities on both species. The biomass of each species was averaged and plotted against the fishing efforts to show contours of biomass. 

The functions can be found in Functions/PopulationEquations.R, Functions/simulateData.R, and Functions/TradeOffAnalysis.R. The functions were plotted through a function in Functions/PlottingThings.Rmd.
```{r}
#plots of Perretti model data
plotTradeOff(Perretti_sim)
```
In the first graph, "Trade Off Contour Analysis-Predator Biomass", total predator biomass is plotted against both predator and forage fish fishing mortality. The contours do not change with forage fish fishing mortality, which suggests that there is no relationship between forage fish fishing dynamics and predator biomass. 

The following two graphs demonstrate predator yield and forage fish yield with changing fishing dynamics, and they show similar results as the yields only change with the specified species fishing mortality. 

A limitation with this two-species model is the chance for populations to become negative. The Perretti model determines population size before coupling and fishing effects with the following equation, where x0 is the initial population size, r is intrinsic growth rate, and K is carrying capacity:

x0*r*(1-(x0/K)

If K is less than x0, the population becomes negative. Because this is not possible in the real world, we chose to abandon the Perretti model and search for a more realistic one. 

